\vspace{-1em}Guidance, Please! Towards a Framework for RDF-based Constraint Languages.

### 2015-07-27 paper/paper 2/dc-2015.tex
paper finalized

The \emph{constraining element} is an intuitive term which indicates the actual type of constraint. For the majority of the constraint types, there is exactly one constraining element, for instance \emph{property domain (R-25, R-26)} restricts domains of properties and there is only one constraining element with exactly the same identifier \emph{property domain}. Complex constraints, however, need several constraining elements to be expressed, for instance \emph{language tag cardinality (R-48, R-49)}\ke{ich hab das etwas umgestellt, bin mir aber nicht sicher, ob r-48/49 complex sind. Ggf. bitte anpassen, eventuell den satz oder absatz auch rausziehen vor "simple constraints", wenn da erst mal geklärt wird, was das ist.} is used to restrict data properties to have a minimum, maximum, or exact number of relationships to literals with selected language tags. Thus, three constraining elements are needed to express each possible constraint of that constraint type. This example also illustrates that the granularity of the constraint types varies and certainly often is debatable. Keep in mind that they correspond to requirements as identified by the working groups. The constraining elements, as in this example, are closer to atomic elements of constraints.

@@

%\emph{Complex constraints} can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar.
%There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
%\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
%These axiom types are sometimes considered as basic axiom types in DL - using some suggestive notation such as
%\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
%Such axioms, however, are just syntactic sugar - 
%all role characteristics can be expressed using the basic features of DL.
%
%The \emph{irreflexive object properties} constraint type (\emph{R-60}) 
%restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
%With the following \emph{irreflexive object property} constraint, for instance, one can state that individuals cannot be authors of themselves:
%\begin{DL}
%\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
%\end{DL}
%When mapped to the \emph{RDF-CV} (see Table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):
%
%%\begin{gcotable}
%%property & $\exists$ marriedTo . Self & marriedTo & - & Self & $\exists$ & - \\
%%class & $\neg$ $\exists$ marriedTo . Self & - & - & $\exists$ marriedTo . Self & $\neg$ & - \\
%%class & $\top$ & - & - & $\top$, $\neg$ $\exists$ marriedTo . Self & $\sqsubseteq$ & - \\
%%\end{gcotable}
%
%\begin{table}[H]
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\caption{Irreflexive Object Properties as Complex Constraints}
	%\label{tab:irreflexive-object-properties-as-complex-constraints}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\exists$ authorOf.Self & marriedTo & - & Self & $\exists$ & - \\
%class & $\neg$ $\exists$ authorOf.Self & - & - & $\exists$ authorOf.Self & $\neg$ & - \\
%class & $\top$ & - & - & $\top$, $\neg$ $\exists$ authorOf.Self & $\sqsubseteq$ & - \\
		%\end{tabular}
%\end{table}
%
%When using the \emph{OWL RBox axiom} \emph{role irreflexivity} in terms of syntactic sugar, 
%the \emph{complex constraint} can be expressed more concisely in form of a \emph{simple property constraint} with exactly the same semantics (see Table \ref{tab:irreflexive-object-properties-as-simple-constraints}):
%
%%\begin{gcotable}
%%property & $\top$ & marriedTo & - & - & irreflexive & - \\
%%\end{gcotable}
%
%\begin{table}[H]
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\caption{Irreflexive Object Properties as Simple Constraints}
	%\label{tab:irreflexive-object-properties-as-simple-constraints}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\top$ & authorOf & - & - & irreflexive & - \\
		%\end{tabular}
%\end{table}
The \emph{primary key properties (R-226)} constraint type is often useful to declare a given (datatype) property as the primary key of a class, so that a system can enforce uniqueness. 
Books, e.g., are uniquely identified by their ISBN, i.e., the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}
which can be represented using the RDF-CV in form of a complex constraint consisting of two simple constraints (see Table \ref{tab:primary-key-properties-as-complex-constraints}).

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
$\top$ & isbn$^{-}$ & isbn$^{-}$ & - & inverse property & - \\
Book & isbn$^{-}$ & - & - & functional property & - \\
		\end{tabular}
\end{table}
\ke{eins der isbn- in der tabelle ist doch falsch, oder? inverse von isbn- ist doch isbn...}

>>>>

The \emph{constraining element} is an intuitive term which indicates the actual type of constraint. For the majority of the constraint types, there is exactly one constraining element, for instance \emph{property domain (R-25, R-26)} restricts domains of properties and there is only one constraining element with exactly the same identifier \emph{property domain}. Some constraint types, however, need several constraining elements to be expressed, for instance \emph{language tag cardinality (R-48, R-49)} is used to restrict data properties to have a minimum, maximum, or exact number of relationships to literals with selected language tags. Thus, three constraining elements are needed to express each possible constraint of that constraint type. This example also illustrates that the granularity of the constraint types varies and certainly often is debatable. Keep in mind that they correspond to requirements as identified by the working groups. The constraining elements, as in this example, are closer to atomic elements of constraints.

@@

%\emph{Complex constraints} can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar.
%There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
%\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
%These axiom types are sometimes considered as basic axiom types in DL - using some suggestive notation such as
%\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
%Such axioms, however, are just syntactic sugar - 
%all role characteristics can be expressed using the basic features of DL.
%
%The \emph{irreflexive object properties} constraint type (\emph{R-60}) 
%restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
%With the following \emph{irreflexive object property} constraint, for instance, one can state that individuals cannot be authors of themselves:
%\begin{DL}
%\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
%\end{DL}
%When mapped to the \emph{RDF-CV} (see Table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):
%
%%\begin{gcotable}
%%property & $\exists$ marriedTo . Self & marriedTo & - & Self & $\exists$ & - \\
%%class & $\neg$ $\exists$ marriedTo . Self & - & - & $\exists$ marriedTo . Self & $\neg$ & - \\
%%class & $\top$ & - & - & $\top$, $\neg$ $\exists$ marriedTo . Self & $\sqsubseteq$ & - \\
%%\end{gcotable}
%
%\begin{table}[H]
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\caption{Irreflexive Object Properties as Complex Constraints}
	%\label{tab:irreflexive-object-properties-as-complex-constraints}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\exists$ authorOf.Self & marriedTo & - & Self & $\exists$ & - \\
%class & $\neg$ $\exists$ authorOf.Self & - & - & $\exists$ authorOf.Self & $\neg$ & - \\
%class & $\top$ & - & - & $\top$, $\neg$ $\exists$ authorOf.Self & $\sqsubseteq$ & - \\
		%\end{tabular}
%\end{table}
%
%When using the \emph{OWL RBox axiom} \emph{role irreflexivity} in terms of syntactic sugar, 
%the \emph{complex constraint} can be expressed more concisely in form of a \emph{simple property constraint} with exactly the same semantics (see Table \ref{tab:irreflexive-object-properties-as-simple-constraints}):
%
%%\begin{gcotable}
%%property & $\top$ & marriedTo & - & - & irreflexive & - \\
%%\end{gcotable}
%
%\begin{table}[H]
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\caption{Irreflexive Object Properties as Simple Constraints}
	%\label{tab:irreflexive-object-properties-as-simple-constraints}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\top$ & authorOf & - & - & irreflexive & - \\
		%\end{tabular}
%\end{table}
The \emph{primary key properties (R-226)} constraint type is often useful to declare a given (datatype) property as the primary key of a class, so that a system can enforce uniqueness. 
Books, e.g., are uniquely identified by their ISBN, i.e., the property \emph{isbn} is inverse functional {\small\ms{$(\ms{funct } isbn\sp{\overline{\ }})$}}
which can be represented using the RDF-CV in form of a complex constraint consisting of two simple constraints (see Table \ref{tab:primary-key-properties-as-complex-constraints}). The meaning of these simple constraints is that ISBN identifiers can only have {\small\ms{$isbn\sp{\overline{\ }}$}} relations to at most one distinct book.

@@

%\begin{table}[H]
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\caption{Primary Key Properties as Complex Constraints}
	%\label{tab:primary-key-properties-as-complex-constraints}
	%\centering
		%\begin{tabular}{c|c|c|c|c|c}
      %\textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%$\top$ & isbn$^{-}$ & isbn & - & inverse property & - \\
%Book & isbn$^{-}$ & - & - & functional property & - \\
		%\end{tabular}
%\end{table}

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
$\top$ & isbn$^{-}$ & isbn & - & inverse property & - \\
Book & isbn$^{-}$ & - & - & $\leq$ & 1 \\
		\end{tabular}
\end{table}
